Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-03-29 paper/main.tex
alex done

@@

% ---------------- C++ table UM Only--------------------------
\begin{table}[hbtp]
\centering
\caption{Regex vs. Comby for C++ (top) and Java (bottom)}
\label{tab:table_cpp1java}
\begin{tabularx}{\columnwidth}{X|r|cr|cr}
\toprule\toprule
\mc{6}{c}{\textbf{C++}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
ompl & 525 & $\frac{595}{2533}$ & 23.49\% &  $\frac{564}{2127}$  & 25.94\%   \\[1ex]
libgraphqlparser& 137 & $\frac{157}{741}$ & 21.19\%  & $\frac{131}{440}$ & 29.77\% \\[1ex]
rethinkdb\_rebirth &471 &  $\frac{545}{2246}$ & 24.27\% & $\frac{492}{2009}$ & 24.49\% \\[1ex]
xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\[1ex]
tiny-diff.-simulator &  195 & $\frac{747}{1282}$ &58.27\% &$\frac{770}{1214}$ & 63.43\% \\[1ex]
\bottomrule
\end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrr}
                  & &       & Valid  & Invalid  & \% \\\midrule
 \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
%                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                          & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
  & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
%                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
                         & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
 \end{tabularx}
\begin{tabularx}{\columnwidth}{X|r|cr|cr}
\toprule
\mc{6}{c}{\textbf{Java}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
 pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
 spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
 javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
 coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
 egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
 apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]\midrule
\end{tabularx}
\begin{tabularx}{\columnwidth}{Xllrrr}
              & & & Valid  & Invalid  & \% \\\midrule
\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
%              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
%              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
\end{tabularx}
\end{table}

@@

\iffalse
\begin{table}[hbtp]
\centering
\caption{Regex vs. Comby for Python++, including redundant mutants.}
\label{tab:table_pyrust}
\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
\toprule\toprule
\mc{6}{c}{\textbf{Python++}} \\\midrule
                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
\textbf{Project} & \textbf{LOC}    & redundant & $\frac{valid}{total}$ & \textbf{\%} & redundant & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
dtw              & 118             & 567       & $\frac{973}{2106}$    & 46.20\%     & 639       & $\frac{866}{1836}$    & 47.19\% & \\[1ex]
notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$    & 18.89\%     & 122       & $\frac{196}{1019}$    & 19.23\% & \\[1ex]
{\makecell{atlassian-python \\ api}} & 236 & $\frac{302}{1304}$  & 207  &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
       & \multicolumn{2}{c|}{Comby} & 333  & 635  & 351  & 25.25\% & \\ \hline
ESD    & 698 & $\frac{1859}{4626}$  & 1642  & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
       & \multicolumn{2}{c|}{Comby} & 1712  & 1548  &	614  & 44.18\% & \\ \hline
fiber  & 358 & $\frac{493}{1402}$  & 163  & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
       & \multicolumn{2}{c|}{Comby} & 458  & 767 	& 241  & 31.24\% & \\ \hline

\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}

 & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
\end{tabularx}
\end{table}
\fi
% ---------------- Python table UM Only --------------------------

@@

% ---------------- Rust table UM Only --------------------------
\begin{table}[hbtp]
\centering
{\small
\caption{Rust (Regex vs. Comby)}
\label{tab:table_rust1}
\begin{tabularx}{\columnwidth}{X|r|cr|cr}
\toprule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
cargo release    & ...    & & & \\[1ex]
\end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrr}
               & & & Valid  & Invalid  & \% \\\midrule
 \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &   &  & \\
               &   & Std.Dev. & &  &   \\\cline{2-6}
 & \mr{3}{Comby} & Mean     &  &  &  \\
 %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
               &   & Std.Dev. & &  &   \\\bottomrule
\end{tabularx}}
\end{table}

\begin{table}[hbtp]
\centering
\caption{Rust (Regex vs. Comby)}
\label{tab:table_rust1}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{|c|c|c|r|r|r|c|}
\hline
 & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline

 & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
\end{tabular}

@@

The mean and median percentage of mutants that are valid is higher for
Comby for all languages.  On average, even for a single modestly sized
file, using parser combinator combinators saves 100-600 failed
compilations.

@@

We estimated mutant equivalency when possible by examining a sample of
10 randomly selected unkilled mutants for each file.  For some tools,
human inspection of the individual unkilled mutants was not feasible.
However, we were of course able to estimtate the percent of unkilled
mutants that were unkillable for our two modes for all files, and
these numbers can be put in the perspective of rates reported in the
literature.  Reports of equivalent mutant percentages in the
literature sometimes report the percent of \emph{all} mutants
(including killed mutants, which obviously cannot be equivalent) that
are equivalent.  We follow Schuler and Zeller~\cite{EquivMut} in
reporting the percentage of \emph{unkilled} mutants that were found to
be equivalent.  Schuler and Zeller found about 45\% of unkilled
mutants of Java programs to be equivalent.  Other reports of lower
rates, e.g, 9\%~\cite{offutt1997automatically} or 7\%~\cite{TCE}, report the percent of
\emph{all} mutants, and so translate to much higher rates of
equivalency for unkilled mutants.  We therefore suggest that as a
fraction of unkilled mutants, a rate of 30\% or fewer equivalent
mutants is good, and a rate below 10\% is very good.

@@

Previous studies comparing mutation tools for Java
\cite{MajorPIT,gopinath2017does,CommACMJavaTool} found that results
for individual projects often varied considerably across tools without
indicating a clear advantage for one tool over another.  Our primary
criteria for comparing tools is whether there is an overall pattern of
clearly less-effective mutation for our approach.  Below, we discuss
the question for each language separately.  Each section also
addresses briefly the question of whether Regex and Comby modes
perform differently in terms of mutation score and equivalent mutants,
which is relevant to RQ1.

@@

We found only one tool that allowed single-file mutation and worked
reliably for C++.  Dextool generally produced fewer mutants than
either mode of our tool, and for some projects produced an extremely
small number of mutants.  For example, for the {\tt
  tiny-differentiable-simulator} project, Dextool only produced a
total of 61 mutants, and gave a much higher mutation score than our
two modes.  Our tools produced more than 600 additional unkilled
mutants, and none of the inspected mutants were equivalent.
Similarly, Dextool only found 2 unkilled mutants for the {\tt
  libgraphqlparser} project.  We doubt that the set of test weaknesses
in the code is adequately represented by such a small number of
mutants when our tools were able to produce significantly more
unkilled mutants.  Even assuming Dextool's mutants are never
equivalent (highly unlikely) and supposing that a third of our
generated mutants are equivalent, it seems likely that Dextool is
missing important mutants in some cases.  Overall, mean mutation
scores were similar for Dextool and both modes of our tool, while
median mutation scores were much lower for our tools, which produced
many more unkilled mutants for some projects.  For C++, there seems to
be no reason to suspect that the universal approach is producing
sub-par mutation testing.

Comparing Regex and Comby modes, mutation scores are very similar, as
are estimated equivalency rates, though Comby produces a somewhat
larger number and percentage of equivalent mutants.  Both modes
produce relatively low mean percentages of equivalent mutants.

@@

For Java, we compare to PIT, arguably the standard for mutation
testing tools in real-world usage (and a tool very frequently used in
the literature) as well as LittleDarwin, a source-based tool.  Our
tool produces lower but comparable mutation scores, and usually
acceptable or excellent equivalency rates.  LittleDarwin produces
\emph{no} equivalent mutants, but is a highly conservative tool, only
implementing a small set of operators, and not including statement
deletion.  Our tools always produced more non-equivalent unkilled
mutants than PIT by a large margin.

Mutation scores for both modes of our tool were again very similar.
For Java, Comby shows a significant advantage in terms of equivalent
mutants, with very low equivalency rates compared to Regex.

@@

Our comparison of Python mutation tools showed the largest degree of
similarity between all tools.  Muation scores and equivalency rates
were all, given the large variance expected from previous studies of
Java tools, quite similar.  Again, there is no clear disadvantage to
using our approach, and a large advantage in terms of development and
maintenance ease.  The similarity extends to the two modes of our
tool, though Comby had a somewhat higher equivalency rate than the
other tools, including the Regexp mode of our tool.  We discuss
possible reasons for this below.

@@

Finally, for Rust, we observed that the Cargo-Mutants tool simply does
not generate many mutants.  Mutation scores were broadly similar to
ours, though lower, and none of the generated mutants we examined were
equivalent, but for one project Cargo-Mutants did not produce
\emph{any} mutants even though our tool was able to produce a large
number of non-equivalent unkilled mutants.  Again, behavior of the two
modes of our tool was similar, with both having acceptable equivalency
rates and similar mutation scores.

@@

Our tools produce somewhat lower mutation scores than other tools, and
higher (though still generally good) equivalency rates than some
tools, but clearly fall into the same broad similarity of behavior
reported for high-quality mutation tools in the literature comparing
Java mutation tools.   The tools with very low or zero equivalency rates, furthermore,
are very conservative in mutation generation, and often simply don't
produce many mutants.  In fact, inspection of the non-comment source
lines as measured by CLOC shown in Tables~\ref{tab:table_java1} and
\ref{tab:table_java2} and Tables~\ref{tab:table_rust1} and
\ref{tab:table_rust2} shows that for many programs, LittleDarwin and
Cargo-Mutants generated \emph{much less than one mutant per actual
  line code}.  Such a conservative approach is likely to produce few
or no actionable (unkilled and non-equivalent) mutants for a program
with a very good, high-coverage test suite.  But such suites are
arguably the ones where mutation testing offers the most promise, by
allowing the detection of very subtle remaining holes in tests for
high-criticality software.

We speculate that our somewhat lower scores and possibly higher
equivalenc rates derive from the simple fact that we generate more
mutants than most tools.  The ease of defining mutation operators,
plus our interest in generating large numbers of mutants and
prioritizing the most interesting, to make mutation testing useful for
projects with extremely high quality test suites, or even using formal
verification, has encouraged a ``let a thousand flowers bloom''
approach to choosing mutation operators.  For example, our approach
provides many more control-flow mutations than most tools, due to our
observation that test suites sometimes fail to check for loop
execution counts other than one and zero.  Few other tools introduce
{\tt break} and {\tt continue} statements into programs, or transform
one into the other.  These mutants are often equivalent, but also
sometimes expose subtle weaknesses in a test suite, in our
experience.  In general, we take our philosophy of mutation from the
proposal by Gopinath et al.~\cite{Limits,gopinath2017mutation} that
expanding the set of mutation operations and classes of such
operations, in the long run, is more important than using only a small
set of ``good'' operators, in that the end-goal of mutation testing is
to identify weaknesses in test suites, not simply to estimate a single score.

The price paid for this
expansiveness in allowing ways to detect weaknessess in testing is a
larger number of mutants to execute, and a somewhat larger number of
equivalent mutants.  However, for our tool, this observation is not a
fundamental limitation or forced choice.  We propose as future work to
split the operators for each level of the hierarchy into two sets, one
set of core operators similar to (but larger than) those provided by
LittleDarwin or Cargo-Mutants, and a set of ``aggressive'' operators
encompassing our full set of operators, a very simple matter due to
our approach to mutant generation.

>>>>

It is true that this approach to mutant generation does pay a cost in
terms of generating invalid mutants that must be rejected at the
compilation stage of the mutation testing process.  However, as
observed by Pizzoleto et al. in their systematic review of mutation
testing~\cite{pizzoleto2019systematic} cost reduction approaches, mutation testing consists of four
stages: 1) execution of a test suite on the un-mutated program, 2)
generation of mutants, 3) execution of the test suite on the mutants,
and 4) analysis (by a human, usually) of the mutants.  It is only steps 3
and 4 that are ``very costly,'' as Pizzoletto et al. note.  Our approach increases the costs of
the second step only, and, as we discuss below, can help reduce
costs of steps 3 and 4.

@@

% ---------------- C++ table UM Only--------------------------
\begin{table}[hbtp]
\centering
\caption{Regex vs. Comby for C++ (top) and Java (bottom)}
\label{tab:table_cpp1java}
\begin{tabularx}{\columnwidth}{X|r|cr|cr}
\toprule
\mc{6}{c}{\textbf{C++}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
ompl & 525 & $\frac{595}{2533}$ & 23.49\% &  $\frac{564}{2127}$  & 25.94\%   \\[1ex]
libgraphqlparser& 137 & $\frac{157}{741}$ & 21.19\%  & $\frac{131}{440}$ & 29.77\% \\[1ex]
rethinkdb\_rebirth &471 &  $\frac{545}{2246}$ & 24.27\% & $\frac{492}{2009}$ & 24.49\% \\[1ex]
xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\[1ex]
tiny-diff.-simulator &  195 & $\frac{747}{1282}$ &58.27\% &$\frac{770}{1214}$ & 63.43\% \\[1ex]
\bottomrule
\end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrr}
                  & &       & Valid  & Invalid  & \% \\\midrule
 \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                          & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
  & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
                         & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
 \end{tabularx}
\begin{tabularx}{\columnwidth}{X|r|cr|cr}
\toprule
\mc{6}{c}{\textbf{Java}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
 pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
 spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
 javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
 coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
 egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
 apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]\midrule
\end{tabularx}
\begin{tabularx}{\columnwidth}{Xllrrr}
              & & & Valid  & Invalid  & \% \\\midrule
\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
\end{tabularx}
\end{table}

@@

% ---------------- Python table UM Only --------------------------

@@

@@

The mean (and median) percentage of mutants that are valid is higher for
Comby for all languages.  On average, even for a single modestly sized
file, using parser combinator combinators saves 100-600 failed
compilations.

@@

We estimated mutant equivalence when possible by examining a sample of
10 randomly selected unkilled mutants for each file.  For some tools,
human inspection of the individual unkilled mutants was not feasible.
However, we were of course able to estimtate the percent of unkilled
mutants that were unkillable for our two modes for all files, and
these numbers can be put in the perspective of rates reported in the
literature.  Reports of equivalent mutant percentages in the
literature sometimes report the percent of \emph{all} mutants
(including killed mutants, which obviously cannot be equivalent) that
are equivalent.  We follow Schuler and Zeller~\cite{EquivMut} in
reporting the percentage of \emph{unkilled} mutants that were found to
be equivalent.  Schuler and Zeller found about 45\% of unkilled
mutants of Java programs to be equivalent.  Other reports of lower
rates, e.g, 9\%~\cite{offutt1997automatically} or 7\%~\cite{TCE}, report the percent of
\emph{all} mutants, and so translate to much higher rates of
equivalence for unkilled mutants.  We therefore suggest that as a
fraction of unkilled mutants, a rate of 30\% or fewer equivalent
mutants is good, and a rate below 10\% is very good.

@@

Previous studies comparing mutation tools for Java
\cite{MajorPIT,gopinath2017does,CommACMJavaTool} found that results
for individual projects often varied considerably across tools without
indicating a clear advantage for one tool over another.  Our primary
criteria for comparing tools is whether there is an overall pattern of
clearly less-effective mutation for our approach.  We only report
results for Comby, as Comby and Regexp approaches produced nearly
identical mutation scores, and had similar equivalence rates.  In
fact, Regexp equivalence rates were somewhat lower over all than those
shown here.  Full results for Regexp mutation execution are included
in our replication package.

@@

We found only one tool that allowed single-file mutation and worked
reliably for C++.  Dextool generally produced fewer mutants than
either mode of our tool, and for some projects produced an extremely
small number of mutants.  For example, for the {\tt
  tiny-differentiable-simulator} project, Dextool only produced a
total of 61 mutants, and gave a much higher mutation score than our
two modes.  Our tools produced more than 600 additional unkilled
mutants, and none of the inspected mutants were equivalent.
Similarly, Dextool only found 2 unkilled mutants for the {\tt
  libgraphqlparser} project.  We doubt that the set of test weaknesses
in the code is adequately represented by such a small number of
mutants when our tools were able to produce significantly more
unkilled mutants.  Even assuming Dextool's mutants are never
equivalent (highly unlikely) and supposing that a third of our
generated mutants are equivalent, it seems likely that Dextool is
missing important mutants in some cases.  Overall, mean mutation
scores were similar for Dextool and our tool, while
median mutation scores were much lower for our tool, which produced
many more unkilled mutants for some projects.  While our equivalence
rate was somewhat high for one project, it was otherwise very good
compared to results in the literature.  For C++, there seems to
be no reason to suspect that the universal approach is producing
sub-par mutation testing.

@@

For Java, we compare to PIT, arguably the standard for mutation
testing tools in real-world usage (and a tool very frequently used in
the literature) as well as LittleDarwin, a source-based tool.  Our
tool produces lower but comparable mutation scores, and usually
acceptable or excellent equivalence rates.  LittleDarwin produces
\emph{no} equivalent mutants, but is a highly conservative tool, only
implementing a small set of operators, and not including statement
deletion.  Our tools always produced more non-equivalent unkilled
mutants than PIT by a large margin.  Our equivalence rates were
uniformly low, with a median of \emph{no} equivalent mutants.

@@

Our comparison of Python mutation tools showed the largest degree of
similarity between all tools.  Muation scores and equivalence rates
were all, given the large variance expected from previous studies of
Java tools, quite similar.  Again, there is no clear disadvantage to
using our approach, and a large advantage in terms of development and
maintenance ease.  Comby had a somewhat higher equivalence rate than the
other tools (including the Regexp mode of our tool).  We discuss
possible reasons for this below.

@@

Finally, for Rust, we observed that the Cargo-Mutants tool simply does
not generate many mutants.  Mutation scores were broadly similar to
ours, though lower, and none of the generated mutants we examined were
equivalent, but for one project Cargo-Mutants did not produce
\emph{any} mutants even though our tool was able to produce a large
number of non-equivalent unkilled mutants.  Our tool did produce a
very large percentage of equivalent mutants for one project, and high
percentage for another project, but these numbers are still well
within the range of expected results in past inspections of
equivalent mutants.

@@

Our tools produce somewhat lower mutation scores than other tools, and
higher (though still generally good) equivalence rates than some
tools, but clearly fall into the same broad similarity of behavior
reported for high-quality mutation tools in the literature comparing
Java mutation tools.   The tools with very low or zero equivalence rates, furthermore,
are very conservative in mutation generation, and often simply don't
produce many mutants.  In fact, inspection of the non-comment source
lines as measured by CLOC shown in Tables~\ref{tab:table_java1} and
\ref{tab:table_java2} and Tables~\ref{tab:table_rust1} and
\ref{tab:table_rust2} shows that for many programs, LittleDarwin and
Cargo-Mutants generated \emph{much less than one mutant per actual
  line code}.  Such a conservative approach is likely to produce few
or no actionable (unkilled and non-equivalent) mutants for a program
with a very good, high-coverage test suite.  But such suites are
arguably the ones where mutation testing offers the most promise, by
allowing the detection of very subtle remaining holes in tests for
high-criticality software.

We speculate that our somewhat lower scores and possibly higher
equivalence rates derive from the simple fact that we generate more
mutants than most tools.  The ease of defining mutation operators,
plus our interest in generating large numbers of mutants and
prioritizing the most interesting, in order to make mutation testing useful for
projects with extremely high quality test suites, or even using formal
verification, has encouraged us to take a ``let a thousand flowers bloom''
approach to choosing mutation operators.  For example, our approach
provides many more control-flow mutations than most tools, due to our
observation that test suites sometimes fail to check for loop
execution counts other than one and zero.  Few other tools introduce
{\tt break} and {\tt continue} statements into programs, or transform
one into the other.  These mutants are often equivalent (in fact, they
give rise to a large number of our equivalent mutants reported above), but \emph{also
sometimes expose subtle weaknesses in a high quality test suite}, in our
experience.  In general, we take our philosophy of mutation from the
proposal by Gopinath et al.~\cite{Limits,gopinath2017mutation} that
expanding the set of mutation operations and classes of such
operations, in the long run, is more important than using only a small
set of ``good'' operators, in that the end-goal of mutation testing is
to identify weaknesses in test suites, not simply to estimate a single
score.

The price paid for this
expansiveness in allowing ways to detect weaknessess in testing is a
larger number of mutants to execute, and a somewhat larger number of
equivalent mutants.  This cost, we observe, falls especially on Comby in some
cases, because Comby is better at identifying places to inject
control flow changes that produce valid code, but do yield equivalent mutants in
many cases.

However, for our approach, this is not a
fundamental limitation or forced choice.  We propose as future work to
split the operators for each level of the hierarchy into (at least) two sets, one
set of core operators similar to (but larger than) those provided by
LittleDarwin or Cargo-Mutants, and a set of ``aggressive'' operators
that includes our full current set of operators, a very simple matter due to
our approach to mutant generation.

In fact, this brings us back to a point made in the introduction.
While our approach introduces some unavoidable costs in the
(relatively cheap, compared to execution and analysis) compilation stage of
mutation testing, it may offer \emph{more} flexibility in addressing
execution and analysis stages.  Defining limited custom sets of
operators to run is trivial in our setting, making it easy to
execute only, e.g., statement mutations, or, in a numeric-computation
intensive codebase, only arithmetic operators.  Expanding the
hierarchical structure to include ``famillies'' of mutation operators
is trivial, and lets users specify the aggressiveness of mutation
generation in a simple way.  Our approach also produces actual source
copies of all generated mutants, making it easy to apply techniques
like Predictive Mutation Testing~\cite{zhangPMT} that work with source
code.  This also makes it easy to integrate, as we have, heuristic
aids to mutant \emph{analysis} such as ranking of unkilled mutants by
similarity and ``interstingness,'' an approach that has already proven
useful in using mutation testing to improve static analysis tools~\cite{StatMut}.
